
\preClass{Basic Expressions}


\begin{problem}
\item Evaluate the following expressions.

  \begin{subproblem}
  \item 
    \begin{eqnarray}
      y & = 3x + 4.
    \end{eqnarray}
    \vfill
  \end{subproblem}


\end{problem}


\actTitle{Probability: Introduction} 

\textit{You are not expected to finish this during the class
  period. Anything you don't finish is good to practice at home.}


\begin{definition}
  The \textit{probability} of an outcome is the long term proportion
  in which the outcome is observed.
\end{definition}


\begin{example}
  Probabilities reveal long term predictability. If the probability is
  0 then the outcome is extremely unlikely to occur. If the
  probability is 1 then the outcome will almost always occur.
\end{example}
\begin{problem}

\item Define the following terms in your own words : (1 or 2 sentences
  each)

  \begin{subproblem}
    \item Event:
      \vfill
    \item Outcome:
      \vfill
    \item Experiment:
      \vfill
    \item Observational Study:
      \vfill
    \item Expected Outcome:
      \vfill
    \item Variation:
      \vfill
  \end{subproblem}

\item \textbf{Experiment}: Flip a coin 10 times and record the total
  sample proportion of heads each time, including the previous flips.

  \begin{tabular}{| c | c | r |}
    \hline 
    Number of Flips & Number of Heads & Sample Proportion of Heads \\ \hline 
    5 &  & \\  [12pt] \hline 
    10 &  & \\  [12pt] \hline 
    15 & & \\  [12pt]  \hline 
    20 & & \\ [12pt]  \hline 
  \end{tabular}  

  \clearpage

\item Find a partner and put your data together. Describe what happens
  to the proportion when you have 10 flips? 20 flips? More?

  \begin{tabular}{| c | c | r |}
    \hline 
    Number of Flips & Number of Heads & Sample Proportion of Heads \\ \hline 
    10 &  & \\  [12pt] \hline 
    20 &  & \\  [12pt] \hline 
    30 & & \\  [12pt]  \hline 
    40 & & \\ [12pt]  \hline 
  \end{tabular}  



\item Make a bar graph of the total frequency of heads and tails for
  your 20 flips and a bar graph for your combined 40. 

  \vfill

\item Forty properties are chosen at random. If the property's zoning
  status allows commercial use a one is recorded. Otherwise, a zero is
  recorded. Briefly explain how this activity is related to the
  previous coin flipping problem.

  \vspace{10em}
  


\clearpage

  \begin{definition}
    The \textit{sample space} is the collection of all possible
    \textit{events}.
  \end{definition}


  \begin{definition}
    \textit{Venn diagrams} and \textit{tree diagrams} are ways to
    visualize the sample space.
  \end{definition}



\item \textbf{Practice:}

  \begin{subproblem}
  \item You flip a fair coin once. Define the sample space of this
    experiment by writing down all the possible events. Also, express
    it using a Venn diagram and then using a tree diagram.

    \vfill

  \item Repeat the exercise again but for flipping a coin 2 times.

    \vfill
 
  \item Finally, repeat the exercise for 3 flips.

    \vfill

  \end{subproblem}

\clearpage

\item In a class of 30 students, 10 chose \textbf{only} art as an
  elective, 2 chose both, 14 chose music. How many chose neither?
  \textit{Draw the sample space using an appropriate diagram.}

  \vfill

\item In a class of 30 students, 20 are business majors, 8 are biology
  majors, and 2 are undecided. Two students are chosen at random and
  the majors are recorded.  \textit{Draw the sample space using an
    appropriate diagram.}

  \vfill

 \end{problem}

\preClass{Basic Probability}


\begin{problem}
\item Problems with the sum and conditional probability rules.

  \begin{subproblem}
  \item 
    \begin{eqnarray}
      y & = 3x + 4.
    \end{eqnarray}
    \vfill
  \end{subproblem}


\end{problem}



\actTitle{Probability and Sampling}

\begin{problem}

\item \textit{This activity will be done as a class.}

  A volunteer will pick beans out of a bag. Every time a bean is
  picked, record the total number of black beans picked in the middle
  column. (Include previous trials.) In the right column record the
  total sample proportion of black beans picked every time you pick
  another bean.

\item Fill in the following table:

  \begin{tabular}{| c | c | r |}
    \hline 
    Number & Number of Black & Proportion of Black \\  \hline 
    1 &  & \\  [12pt] \hline 
    2 & & \\  [12pt] \hline 
    3 & & \\  [12pt] \hline 
    4 & & \\  [12pt]  \hline 
    5 & & \\  [12pt] \hline 
    6 & & \\  [12pt] \hline 
    7 & & \\  [12pt] \hline 
    8 & & \\  [12pt] \hline 
    9 & & \\  [12pt] \hline 
    10 & & \\  [12pt] \hline 
  \end{tabular}

\item Based on the results what is the predicted proportion of black beans? 


\end{problem}

 
\actTitle{Probability Rules}
 
\begin{problem}
  \item Define the following terms in your own words: (1-2 sentences) 

    \begin{subproblem}
      \item \textit{Mutually exclusive/disjoint}: 
        \vfill
      \item \textit{Addition rule:} 
        \vfill
      \item \textit{Complement Rule:} 
        \vfill
      \item \textit{Multiplication Rule:} 
        \vfill
      \item \textit{Conditional Probabilities:}  
        \vfill
    \end{subproblem}

  \item Determine each of the following probabilities. For each
    problem make a sketch of the sample space using either a Venn
    diagram or a tree diagram.

  \begin{subproblem}

  \item A single card is pulled at random from a well shuffled,
    standard 52 card deck. What is the probability of drawing a king
    or a queen?

    \vfill
    \vfill

    \clearpage

  \item What is the probability of rolling a 1 or a 6 on a fair six
    sided die?

    \vfill

  \item The probability of \textit{S} is 0.6, the probability of
    \textit{Y} is 0.4, the probability of \textit{S} and \textit{Y} is
    0.5. What is the probability of \textit{S} or \textit{Y}?.

    \vfill

  \item A single card is pulled at random from a well shuffled,
    standard 52 card deck. What is the probability of picking a queen
    or a heart?

    \vfill

  \item What is the probability of not rolling a 6 or rolling a 1 on a
    fair, six sided die?

    \vfill

    \clearpage

  \item What is the probability of getting 3 heads in a row when
    flipping a fair coin?

    \vfill

  \item What is the probability of getting a $\{H,T,H\}$ when flipping
    a fair coin three times?

    \vfill

  \item A fair six sided die is rolled, and the number is even.  What
    is the probability the number is a 4?

    \vfill

  \item A single card is drawn from a well shuffled, standard 52 card
    deck of cards, and the card is is red.  What is the probability
    that the card is a queen? (Show all of your work!)

    \vfill

    \clearpage

  \item Two cards are drawn from a well shuffled, standard 52 card
    deck without replacement.  What is the probability that a king was
    drawn followed by a queen?

    \vfill

  \item You roll a fair, six sided die twice. What is the probability
    of getting exactly one 1?

    \vfill

  \item Two cards are drawn from a well shuffled, standard 52 card
    deck without replacement. What is the probability of first picking
    a queen followed by a card that is not a heart?

    \vfill

  \item A fair, six sided die is rolled twice.  What is the
    probability that the largest number is at least 4?

    \vfill

  \end{subproblem}

\end{problem}




