\documentclass[12pt]{article}

\oddsidemargin=0.0in
\evensidemargin=0.0in
\textwidth=6in
\topmargin=-1in
\textheight=9.5in

\setlength{\parindent}{0pt}

\begin{document}

 \textbf{Probability: Introduction} \\
\\

\textbf{Definition:}
\\
 The \textit{probability} of an outcome is the long term proportion in which the outcome is observed. \\
\\
Probabilities reveal long term predictability. \\
\\
If probability = 0, outcome is \textit{impossible}. \\
If probability is = 1, outcome is a \textit{certainty}. \\ 

\textbf{Define the following terms in your own words :} (1 or 2 sentences) \\
Event:\\  [12pt]
Outcome: \\ [12pt]
Experiment:\\ [12pt]
Observational Study: \\ [12pt]
Expected Outcome:\\ [12pt]
Variation:\\

 \textbf{Experiment :}

 Flip a coin 10 times and record the proportion of heads each time.\\

\begin{tabular}{| c | c | r |}
 \hline 
Number of Flips & Number of Heads & Sample Proportion of Heads \\  \hline 
1 &  & \\  [12pt] \hline 
2 & & \\  [12pt]  \hline 
3 & & \\ \hline 
4 & & \\  \hline 
5 & & \\  \hline 
\end{tabular}
\\
Find a partner and put your data together. What happens to the proportion when you have 20 flips? 30 flips? More?\\
\\
20 Flips - Proportion: \\
30 Flips - Proportion: \\
\noindent \textit{ See page 233 in text for graph of data.}

Plot head and tails freq. in bar chart 

\clearpage 

\textbf{Application:} class

Determine the probability of picking a black bean out of the bag. \\

 Pick a bean randomly out of your bag and record the proportion of black beans. \\
\\
\begin{tabular}{| c | c | r |}
 \hline 
Number & Number of Black & Proportion of Black \\  \hline 
1 &  & \\ \hline 
2 & & \\ \hline 
3 & & \\ \hline 
4 & & \\  \hline 
5 & & \\  \hline 
6 & & \\ \hline 
7 & & \\ \hline 
8 & & \\ \hline 
9 & & \\ \hline 
10 & & \\ \hline 
\end{tabular}
\\
\\
Based on the results what is the predicted proportion of black beans? \\

\textbf{Definititon:}\\

\textit{Sample space:}  is the collection of all possible \textit{events}. \\

\textit{Venn Diagrams} and \textit{ Tree Diagrams} are ways to visualize sample space.
\\

\textbf{Practice:}

\begin{enumerate}
\item You flip a coin. What are the events?\\

\begin{itemize}
 \item What if you flip a coin 2 times?\\
 
 \item How about 3 times? Draw a tree diagram to represent all possible events when you flip a coin 3 times.\\
\vspace{5em}

  \end{itemize}
  
\item In a class of 30 students, 10 chose art as an elective and 14 chose music. How many chose neither?\\
Draw the sample space in a diagram.\\ fix , venn  or tree

 \end{enumerate}
 
 \clearpage
 
\textbf{Definititons:}\\

\textit{Mutually exclusive/disjoint}:\\
\textit{Addition rule:}\\
\textit{Complement Rule:}\\
\textit{Multiplication Rule:}\\
\textit{Conditional Probabilities:}\\

\textbf{Examples:}\\

\begin{enumerate}
\item What is the probability of drawing a king and a queen from a standard 52 card deck?\\
\\
\item What is the probability of drawing a king or a queen from a standard 52 card deck?\\ 
\\
\item What is the probability of rolling a 1 or a 6 on a standard six sided die?\\
\\
\item You roll two dice. What is the probability of getting exactly one 1. \\
\\
\item The probability of \textit{S} is 0.6, the probability of \textit{Y} is 0.4 the probability of \textit{S} and \textit{Y} is 0.5. What is the probability of \textit{S} of \textit{Y}?. \\  

\item What is the probability of picking a queen or a heart?\\

\item What is the probability of picking a queen and not a heart?\\

\item What is the probability of not rolling a 6 or a 1 ? 

\item What is the probability of getting 3 heads when flipping a coin?\\

\item What is the probability of getting a H T H when flipping a coin? \\

\item What is the probability of rolling at least a 4?\\

\item What is the probability that an outcome of rolling a die is 4 given that the number is even?\\

\item What is the probability of picking a queen out a deck of cards given that the card is red?\\


\end{enumerate}



\end{document}




