\preClass{Calculating Sample Means}


\begin{problem}
\item Examples of calculating sample means.

  \begin{subproblem}
  \item 
    \begin{eqnarray}
      y & = 3x + 4.
    \end{eqnarray}
    \vfill
  \end{subproblem}


\end{problem}


\actTitle{Calculating Sample Means and Sample Standard Deviations} 


\begin{problem}

\item The following questions refer to the following set of data: \\
  \begin{tabular}{rrrrrrrr}
    9 & 14.3 & 11.2 & 12.7 & 16.8 & 12.8 & 9.7 & 23.2
  \end{tabular}

  \begin{subproblem}
  \item Make a sketch of a strip chart for the data. Provide a brief
    description about the data and note any patterns.

    \vspace{8em}


  \item Calculate the sample mean, sample standard deviation, and the
    coefficient of variation for the data.

    \vfill

  \item Remove the largest number and then calculate the sample mean,
    the sample standard deviation, and the coefficient of variation
    for the data.

    \vfill

  \item Calculate the percent difference in the values of the sample
    mean and sample standard deviations.

    \vspace{4em}

  \end{subproblem}

  \clearpage

\item A random collection of firms that have filed for chapter 7 bankruptcy will
  be polled and asked how long they have been in bankruptcy. Your
  group believes that the mean length of time in months is 2.3 months
  with a standard deviation of 0.6 months.

  \begin{subproblem}
  \item Suppose that $X$ is a random variable with mean $\mu$ and
    standard devitation $\sigma$. Chebychev's inequality is
    \begin{eqnarray*}
      p\left( -k \sigma \leq X - \mu \leq k \sigma \right) \leq \frac{1}{k^2}.
    \end{eqnarray*}
    Your group plans on polling one-hundred firms. How many of the
    times do you expect to be between 2.0 months and 2.6 months.

    \vfill

  \item Your group plans on polling one-hundred firms. How many of the
    times do you expect to be between 1.1 months and 3.5 months.

    \vfill

  \item Your group plans on polling one-hundred firms. How many of the
    times do you expect to be between 0.8 months and 3.8 months.

    \vfill

    \clearpage

  \item Your group takes an initial poll, and the results are given
    below. Determine the sample mean and sample standard deviation of
    the sample. \\
    \begin{tabular}{rrrrrrrr}
      3.1 & 1.8 & 1.6 & 3.0 & 3.0 & 2.7 & 3.2 & 2.7
    \end{tabular}

    \vfill

  \item Your group takes an initial poll, and the results are given
    below. Determine the sample mean and sample standard deviation of
    the sample. \\
    \begin{tabular}{rrrrrrrr}
      2.6 & 2.4 & 3.0 & 2.7 & 1.9 & 2.3 & 2.8 & 2.7
    \end{tabular}

    \vfill

  \item Why are the results different?

    \vspace{4em}

  \end{subproblem}

\end{problem}

\actTitle{Calculating The Five Point Summary and Box Plots} 

\begin{problem}

\item We will re-examine the data used in the first problem of the
  previous activity. The following questions refer to the following set of data: \\
  \begin{tabular}{rrrrrrrr}
    9 & 14.3 & 11.2 & 12.7 & 16.8 & 12.8 & 9.7 & 23.2
  \end{tabular}

  \begin{subproblem}
  \item Calculate the five point summary for the data and sketch a boxplot.

    \vfill

  \item Remove the largest number and then calculate the five point
    summary for the new data set and sketch a boxplot.

    \vfill

  \item Calculate the percent difference in the values of the sample
    median and the sample quartiles.

    \vspace{4em}

  \item Compare the percent differences for the sample mean and the
    sample median. Which one is smaller?

    \vspace{4em}

  \end{subproblem}

  \clearpage

\item The femur lengths for a randomly chosen group of people are
  measured. The length in millimeters of each person's femur and their
  sex are given in the following table: \\
  \begin{tabular}{rr|rr|rr|rr|rr}
    413.2 & F & 411.1 & F & 431.6 & F & 365.6 & F & 450.8 & M \\
    456.5 & F & 415.7 & F & 407.3 & F & 430.4 & M & 434.7 & F \\
    435.9 & M & 479.4 & M & 451.3 & M & 426.3 & F & 427.0 & M \\
    432.4 & M
  \end{tabular}

  \begin{subproblem}
  \item Determine the sample mean, sample standard deviation, and the
    five point summary for the females in the group.

    \vfill

  \item Determine the sample mean, sample standard deviation, and the
    five point summary for the males in the group.

    \vfill

  \item Use the previous information to estimate the mean and
    standard deviation of the group as a whole. 

    \vfill

  \item The hero on your favorite television show finds a skeleton and
    measures the femur on the skeleton. The length of the femur is
    421.0 mm. What is your best guess for the gender? How sure are you
    and why?

    \vspace{4em}

  \end{subproblem}

\clearpage

\item Calculate a weighted mean. Suppose that the grade for a class is
  calculated using the following weights: \\
  \begin{tabular}{rr}
    60\% & Exam Mean \\
    20\% & Quiz Mean \\
    20\% & Attendance 
  \end{tabular}

  Suppose that a student has a 90\% mean quiz grade, a 95\% attendance
  mean, and the two tests are an 80\% and a 75\%. 

  \begin{subproblem}
  \item Determine the student's current weighted grade.

    \vfill

  \item Suppose that the student has one more test to take. What grade
    does the student need in order to get an 87\% in the course?

    \vfill

  \end{subproblem}


\end{problem}

