
\lecture{Geometric and Hypergeometric Distributions}{geometry-hypergeometric}
\section{Geometric and Hypergeometric Distributions}

\title{Geometric and Hypergeometric Distributions}
\subtitle{Don't Stop Me Now!}

%\author{Kelly Black}
%\institute{Clarkson University}
\date{23 September 2013}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


\iftoggle{clicker}{%
  \begin{frame}
    \frametitle{Clicker Quiz}

    A fair die is rolled ten times. What is the probability of getting
    4 ones?

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: ${10 \choose 6} \lp\frac{1}{6}\rp^6 \lp\frac{5}{6}\rp^4$ & 
      B: ${10 \choose 4} \lp\frac{1}{6}\rp^4 \lp\frac{5}{6}\rp^6$  \\ [10pt]
      C: ${10 \choose 6} \lp\frac{1}{6}\rp^4 \lp\frac{5}{6}\rp^6$ & 
      D: ${10 \choose 4} \lp\frac{1}{6}\rp^6 \lp\frac{5}{6}\rp^4$
    \end{tabular}

    \vfill
    \vfill
    \vfill


  \end{frame}
}

\subsection{Geometric Distribution}

\begin{frame}
  \frametitle{Geometric Distribution}

  \begin{definition}
    A random variable is a \redText{Geometric distribution} if it
    consists of
    \begin{itemize}
    \item A sequence of Bernoulli Trials, $X_1$, $X_2$, $X_3$,
      $\ldots$
    \item Each Bernoulli trial is independent of the others.
    \item Each one has a probability of $p$ of success.
    \item Test each one in order starting at the first.
    \item You report the number of tests you conduct when you get the
      first ``1'' (i.e. a ``success''). 
    \end{itemize}
  \end{definition}


\end{frame}

\begin{frame}{Properties}

  The probability mass function for a geometric distribution with
  probability $p$ for a success in any one trial is
  \begin{eqnarray*}
    p(X=k) & = & \lp 1-p\rp^{k-1} p.
  \end{eqnarray*}

  The expectation and variance of a geometric distribution are
  \begin{eqnarray*}
    E[X] & = & \frac{1}{p}, \\
    \mathrm{Var}[X] & = & \frac{1-p}{p^2}.
  \end{eqnarray*}

\end{frame}


\begin{frame}{Examples}

  \begin{itemize}
  \item Pick tires at random (with replacement) from a production line
    and test each one until I find one that fails an
    inspection. Report the number of
    tires that I tested.\\
  \item Call people at random (with replacement) and ask if they
    support candidate A. The first time someone says ``yes'' report
    the number of people that I called.
  \item A person goes surfing in shark infested waters. Each time the
    person goes surfing there is a probability of 0.0002 that the
    person will be eaten. Report the number of times that the person
    leaves the beach to go surfing.
  \end{itemize}
  
\end{frame}

\begin{frame}{Example}

  I walk to class every day. On any given day there is a probability
  of 0.01 that I get frost bite. The first time that I get frost bite
  I decide this is enough and enroll at Mississippi State
  University. I go to class every day. What is the probability that I
  enroll at Mississippi State University on the thirty-first day?

\end{frame}

\subsection{Hypergeometric Distribution}

\begin{frame}{Hypergeometric Distribution}

  \begin{itemize}
  \item I have $N$ objects.
  \item There are $r$ identical objects of type I. (Think red marble.)
  \item There are $N-r$ identical objects of type II. (Think blue  marble.)
  \item I choose $n$ objects at random \textbf{without} replacement.
  \item I report the number of objects ($k$) that are type I. (red marbles)
  \end{itemize}

  \vfill


\end{frame}

\begin{frame}{Hypergeometric Distribution}


  \begin{definition}[Hypergeometric Distribution]
    If $X$ is a hypergeometric distribution then
    \begin{eqnarray*}
      p(X=k) & = & \frac{{r \choose k}\cdot{{N-r}\choose{n-k}}}{{N \choose n}}, \\
      E[X] & = & \frac{nr}{N}, \\
      \mathrm{Var}[X] & = & \frac{N-n}{N-1} \cdot n \cdot \frac{r}{N} \cdot \lp 1-\frac{r}{N}\rp.
    \end{eqnarray*}
  \end{definition}

\end{frame}

\subsection{Examples}




\begin{frame}{Example}

  A jury pool has twenty men and fifteen women. A jury of twelve
  people is to be chosen. If everybody is treated equally then what is
  the probability that a jury will have exactly two women on it?

\end{frame}



\iftoggle{clicker}{%
  \begin{frame}{Clicker Quiz}

    \vfill

    A contestant on a TV game show must throw a ball through a hoop
    before moving on to the next task. Player A has a probability of
    0.45 of getting the ball through the hoop on any one throw. What
    is the probability that the player will be able to advance on the
    fourth try?

    \vfill

    What kind of distribution is this?

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: Binomial  & B: Geometric & C: Hypergeometric
    \end{tabular}

    \vfill
  
  \end{frame}
}



\begin{frame}
  \frametitle{Example}

  A contestant on a TV game show must throw a ball through a hoop
  before moving on to the next task. Player A has a probability of
  0.45 of getting the ball through the hoop on any one throw. What is
  the probability that the player will be able to advance on the
  fourth try?

  \vfill

\end{frame}

\iftoggle{clicker}{%
  \begin{frame}{Clicker Quiz}

    \vfill

    A batch of thirty concrete cores is made. Seven of the cores are
    below the standards. Five cones are chosen at random. What is the
    probability that two will be below the standards?

    \vfill

    What kind of distribution is this?

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: Binomial  & B: Geometric & C: Hypergeometric
    \end{tabular}

    \vfill
  
  \end{frame}
}



\begin{frame}
  \frametitle{Example}

    A batch of thirty concrete cores is made. Seven of the cores are
    below the standards. Five cones are chosen at random. What is the
    probability that two will be below the standards?
  

\end{frame}

\iftoggle{clicker}{%
  \begin{frame}{Clicker Quiz}

    \vfill

    I check the load on a circuit every sixty seconds. There is a
    probability of 0.03 that the current exceeds the safety limit
    during any check. What is the probability that the safety limits
    are exceeded on four of those checks?

    \vfill

    What kind of distribution is this?

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: Binomial  & B: Geometric & C: Hypergeometric
    \end{tabular}

    \vfill
  
  \end{frame}
}


\begin{frame}
  \frametitle{Example}

  I check the load on a circuit every sixty seconds. There is a
  probability of 0.03 that the current exceeds the safety limit during
  any check. What is the probability that the safety limits are
  exceeded on four of those checks?

  \vfill

\end{frame}



% LocalWords:  Clarkson pausesection hideallsubsections
%  LocalWords:  Hypergeometric hypergeometric
