
\lecture{Hypothesis Testing Redux}{hypothesis-testing-redux}
\section{Hypothesis Testing Redux}

\title{Hypothesis Testing Again}
\subtitle{Examples - Which one to use?}

%\author{Kelly Black}
%\institute{Clarkson University}
\date{8 November 2013}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


%\begin{frame}
%  \frametitle{Clicker Quiz}%
%
%  \iftoggle{clicker}{%
%
%    I read in the \textit{Integrator} that the mean height of Clarkson
%    students is 180.0 cm with a standard deviation of 14.0 cm. I think
%    that this is too high and sample twenty students chosen at random
%    to find a sample mean of 174.7 cm. Are my suspicions correct? (Use
%    a 5\% significance level.)
%
%    \vfill
%
%    \begin{tabular}{l@{\hspace{3em}}l}
%      A: & There is sufficient evidence to reject $H_0$.  \\
%      B: & There is not sufficient evidence to reject $H_0$.  
%    \end{tabular}
%
%    \vfill
%    \vfill
%    \vfill
%  }
%
%\end{frame}

\begin{frame}
    \input{img/confidenceIntervalDist}  % Added [thick,scale=0.6, every node/.style={scale=0.6}]
                                      % to the \begin{tikzpicture} line
\end{frame}

\subsection{Examples}

\begin{frame}
  \frametitle{What if....}


  I read in the \textit{Integrator} that the mean height of Clarkson
  students is 180.0 cm. I think that this is too high and sample
  twenty students chosen at random to find a sample mean of 174.7 cm
  with a {\color{red}sample} standard deviation of 14 cm. Are my
  suspicions correct?  (Use a 5\% significance level.)

    \vfill
    
    \only<2->{%
      We do not have sufficient evidence to reject $H_0$ at the 5\%
      significance level using a $t$-distribution with twenty samples and
      a standard deviation of 14.0cm.
    }
\end{frame}




\begin{frame}{Example}

  I see a publication that says that the mean value of houses in an
  area is \$285,000. I think that this is incorrect.  Suppose that I
  take a random sample of seventeen homes and find a sample mean of \$
  and a sample standard deviation of \$35,000. Are my suspicions
  correct?  (Use a 5\% significance level.)

  \vfill

  \only<2->{%
    We have sufficient evidence to reject $H_0$ at the 5\%
    significance level assuming a $t$-distribution with sixteen
    degrees of freedom.  }
\end{frame}


\begin{frame}{Example}

    \begin{eqnarray*}
      \begin{array}{lrcl}
        H_0: & \mu & = & 800 \\
        H_a: & \mu & > & 800
      \end{array}
    \end{eqnarray*}

    \begin{eqnarray*}
      n & = & 30, \\
      \bar{x} & = & 758,\\
      \sigma & = & 118, \\
      \alpha & = & 0.05.
    \end{eqnarray*}


    \only<2->{%
      We have sufficient evidence to reject $H_0$ at the 5\%
      significance level using a normal distribution with thirty
      samples and a standard deviation of 118.  }
\end{frame}



\begin{frame}{Example}

  The breaking strength of wood pegs is to be tested. We think that
  the mean breaking strength is 50.0N with a standard deviation of
  8.0N. How many samples should we take if we want the 95\% confidence
  interval to have an error 0.2N?
  
\end{frame}


\begin{frame}{Example}

  The concentration of lead is tested at twelve randomly chosen
  homes. The tests yield a sample mean of 14.5 parts per billion with
  a sample standard deviation of 3.8 parts per billion. Determine the
  95\% confidence interval.
  
\end{frame}

\begin{frame}{Example}

  The EPA limits for lead in tap water is 15 parts per billion.  The
  concentration of lead is tested at twelve randomly chosen homes. The
  tests yield a sample mean of 14.5 parts per billion with a sample
  standard deviation of 3.8 parts per billion. Is the mean lead levels
  below the EPA limits?
  
\end{frame}


% LocalWords:  Clarkson pausesection hideallsubsections
