
\lecture{Intersection of Event}{intersection-events}
\section{Intersection of Events}

\title{Intersection of Events}
\subtitle{Combinations of Events}

%\author{Kelly Black}
%\institute{Clarkson University}
\date{1 September 2014}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


\iftoggle{clicker}{%
  \begin{frame}
    \frametitle{Clicker Quiz}

    I have a bag with ten blue marbles and four red marbles. What is
    the probability that I draw a red first and then a blue marble
    second?

    \vspace{2em}

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: $\frac{12}{182}$ & B: $\frac{12}{196}$ & C: $\frac{40}{182}$ & D: $\frac{40}{196}$
    \end{tabular}

    \vfill

  \end{frame}
}

\subsection{Conditional Probability Examples}

\begin{frame}
  \frametitle{Example}

  A bag has thirty red marbles and eight green marbles. What is the
  probability of pulling three red marbles out in a row without
  replacement?

  \vfill

\end{frame}


\begin{frame}
  \frametitle{Example}

    In a large population of students, 61\% attend a four year
    institution, and the rest attend a two year institution. Of those
    going to a four year institution 52\% are female. Of those going
    to a two year institution 47\% are female. What is the gender
    distribution of students?

    \vfill

\end{frame}

\subsection{Independence}

\begin{frame}
  \frametitle{Independence}

%  \begin{definition}
%    Two events are independent if 
%    \begin{eqnarray*}
%      p(A|B) & = & p(A).
%    \end{eqnarray*}

%    This is equivalent to requiring that
%    \begin{eqnarray*}
%      p(B|A) & = & p(B).
%    \end{eqnarray*}

%  \end{definition}

\end{frame}

\begin{frame}{Example}

  Suppose that $p(A)=0.6$, $p(B)=0.3$, and $p(A\mathrm{~or~}B)=.8$
  Are they independent?

  \vfill

  
\end{frame}



\subsection{Examples}


\iftoggle{clicker}{%
  \begin{frame}
    \frametitle{Clicker Quiz}

    A bag has three red marbles and two green marbles. Event A is the
    event that I draw a red marble first. Event B is that I draw a red
    marble second. Are these two events independent?

    \vspace{2em}

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: Independent & B: Not Independent 
    \end{tabular}

    \vfill

  \end{frame}
}





\begin{frame}{Many Events}

  \begin{eqnarray*}
    p(A_1 \cap A_2 \cap A_3 \cap \cdots \cap A_N) & = & ? \\
    \uncover<2>{
      p(A_1 \cup A_2 \cup A_3) & = & ? \\
    }
  \end{eqnarray*}
  
\end{frame}

