
\lecture{Linear Combinations of Normal Distributions}{normal-distributions-linear-combos}
\section{Linear Combinations of Normal Random Variables}

\title{Linear Combinations of Normal Random Variables}
\subtitle{Adding Normals}

%\author{Kelly Black}
%\institute{Clarkson University}
\date{8 October 2014}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


\iftoggle{clicker}{%
  \begin{frame}{Clicker Quiz}

    What is the probability that a random variable that follows a
    standard normal will be less than -0.45?

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l}
      A:  .3264  & B: .3632  & C: .6736
    \end{tabular}

    \vfill
    \vfill
    \vfill

  \end{frame}
}


\subsection{Linear Combinations}

\begin{frame}{Linear Transformation of a Normal Variable}

  \vfill

  Suppose that $X$ is a normally distributed random variable with mean
  $\mu_x$ and standard deviation $\sigma_x$.

  \vfill

  Define a new random variable
  \begin{eqnarray*}
    Y & = & a X + b.
  \end{eqnarray*}

  \vfill

  Then $Y$ is normally distributed with
  \begin{eqnarray*}
    E[Y] & = & \\
    \mathrm{Var}[Y] & = & 
  \end{eqnarray*}

  \vfill

\end{frame}


\begin{frame}{Sum of Two Normal Variables}

  \vfill

  Suppose that $X$ is a normally distributed random variable with mean
  $\mu_x$ and standard deviation $\sigma_x$. Also, $Y$ is a normally
  distributed random variable with mean $\mu_y$ and standard deviation
  $\sigma_y$. 

  \vfill

  Define a new random variable
  \begin{eqnarray*}
    W & = & X + Y.
  \end{eqnarray*}

  \vfill

  \textit{\redText{If $X$ and $Y$ are independent}} then $W$ is
  normally distributed with
  \begin{eqnarray*}
    E[W] & = & \\
    \mathrm{Var}[W] & = & 
  \end{eqnarray*}

  \vfill
  
\end{frame}

\begin{frame}{Linear Combinations of Normal Variables}
  
  \vfill

  Suppose that $X_1$, $X_2$, \ldots, $X_n$ are \redText{independent}
  normally distributed random variables with mean $\mu_1$, $\mu_2$,
  \dots, $\mu_n$ and standard deviations $\sigma_1$, $\sigma_2$,
  \ldots, $\sigma_n$.

  \vfill

  Define a new random variable
  \begin{eqnarray*}
    W & = & a_1 X_1 + a_2 X_2 + \cdots + a_n X_n + b,
  \end{eqnarray*}
  where $a_1$, $a_2$, \ldots, $a_n$ and $b$ are all constants.

  \vfill

  Then $W$ is normally distributed with
  \begin{eqnarray*}
    E[W] & = & \\
    \mathrm{Var}[W] & = & 
  \end{eqnarray*}

  \vfill
  
\end{frame}


\subsection{Examples}

\begin{frame}
  \frametitle{Example}

  The temperature measurements of a heat sink is normally distributed
  with mean 112F and standard deviation 14F. What is the probability
  that a measurement will be between 47C and 55C?

  \vfill



\end{frame}





\begin{frame}
  \frametitle{Example}

  The mass of glass jars are normally distributed with a mean of
  0.25kg and a standard deviation of 0.03kg. The total mass of the
  packaged product is normally distributed with a mean of 0.35kg and a
  standard deviation of 0.08kg. What is the \redText{distribution}
  associated with the contents of the jars? (Assume that the mass of
  the jar and its contents are independent.)

\end{frame}

\iftoggle{clicker}{%
  \begin{frame}
    \frametitle{Clicker Quiz - Starting Salaries}

    $X$ is normally distributed with mean 3.0 and std. deviation 1.7.
    
    $Y$ is normally distributed with mean 4.1 and std. deviation 2.1.

    $W$ is normally distributed with mean 2.5 and std. deviation 3.1.

    What is the standard deviation of  $X+Y-W$?


    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l}
      A:  0.7  & B: 4.11  & C: 6.90
    \end{tabular}

    \vfill
    \vfill
    \vfill



  \end{frame}
}



\begin{frame}
  \frametitle{Example}

  $X$ is normally distributed with mean 3.0 and std. deviation 1.7.

  $Y$ is normally distributed with mean 4.1 and std. deviation 2.1.

  $W$ is normally distributed with mean 2.5 and std. deviation 3.1.

  What is the probability that $X+Y-W$ is between 3.0 and 4.0?

  \vfill

\end{frame}





\begin{frame}
  \frametitle{Example}

  The width of glass panes is normally distributed with a mean of
  1.20cm and a standard deviation of 0.08cm. What mean width and
  standard deviation should be designed for the groove in the frames
  so that the gap is normally distributed with a mean of 0.15cm and a
  standard deviation of 0.12cm?

  \vfill

  ($\mu=1.35$ cm, $\sigma\approx$.089 cm)

\end{frame}


\begin{frame}
  \frametitle{Example}

  The voltage across a resistor is normally distributed with mean 18V
  and standard deviation 1.2V. Two measurements are taken and
  ``averaged.'' What is the probability that the ``average'' is more
  than 18.5V? (Assume that the measurements are independent.)

  \vfill

  (0.2776)

\end{frame}


\iftoggle{clicker}{%
  \begin{frame}
    \frametitle{Clicker Quiz}

    $X_1$ is normally distributed with mean 80.0 and standard
    deviation 2.0.  $X_2$ is also normally distributed with mean 80.0
    and standard deviation 2.0. What is the standard deviation of
    \begin{eqnarray*}
      \frac{1}{2} X_1 + \frac{1}{2} X_2?
    \end{eqnarray*}

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: 2.0  & B: 1.41 & C: 3.07 & D: 4.0
    \end{tabular}

    \vfill
    \vfill
    \vfill

  \end{frame}
}




\begin{frame}{Example}

  Some bolts are to be designed so that their breaking strength is
  normally distributed with mean 80N. Three samples are taken. What
  should the standard deviation be so that the probability that the
  ``average'' breaking strength of the samples is less than 78N is
  0.05?

  (Assume that the samples are independent.)

  \vfill

  (2.71N)
  
\end{frame}


\begin{frame}{Example}
  
  A random variable, $X$, is normally distributed with mean 3.5 and
  standard deviation 1.2. Determine the following probabilities:
  \begin{eqnarray*}
    p(X \geq 3.2), (.5987)\\
    p(X \leq 4.7), (.1587) \\
    p(3.1 \leq X \leq 5.2),  (.5523)\\
    p(|X-3| \leq 1.1), (.6003)\\
    p(|X-2.5| \geq 2.1), (.1846).
  \end{eqnarray*}
  
\end{frame}

\begin{frame}{Example}

  A cylinder and piston is to be constructed so that the diameter of
  the cylinder is normally distributed with mean 4.50 cm and a
  standard deviation of 0.02cm. The bore should be designed so that
  the gap is normally distributed with a mean of 0.10cm and a standard
  deviation of 0.04cm. What is the required distribution for the bore?

  \vfill
  
  Normal: mean=4.70cm, std. dev$\approx$.077cm
  
\end{frame}

\begin{frame}
  A random variable is normally distributed with mean 0.05 and
  standard deviation 0.35. It is sampled four times (assume
  independence between samples), and the sample mean, $Y$, is
  calculated. What is the probability that the sample mean of the four
  samples is more than 0.10 from the mean of $Y$?

  \vfill

  (.5687)

\end{frame}


\begin{frame}
  A random variable is normally distributed with mean 0.05 and
  standard deviation 0.35. It is sampled five times (assume
  independence between samples), and the sample mean, $Y$, is
  calculated. What is the probability that the sample mean of the five
  samples is more than 0.10 from the mean of $Y$?

  \vfill

  (.5222)
\end{frame}


% LocalWords:  Clarkson pausesection hideallsubsections hideothersubsections
% LocalWords:  sectionstyle
