
\lecture{More Sample Distributions}{more-sample-distributions}
\section{More Sample Distributions}

\title{More Sample Distributions}
\subtitle{More Examples}

\author{Kelly Black}
\institute{Clarkson University}
\date{Feb 27, 2012}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[pausesection,hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


\begin{frame}
  \frametitle{Clicker Quiz}

  The mass of the contents of a product are normally distributed with
  a mean of 0.35 kg and a standard deviation of 0.04 kg. What is the
  probability that the sample mean of four samples will be less than
  0.33 kg?

  \vfill

  \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
    A: 0.0228  & B: 0.1318  & C: 0.1587
  \end{tabular}

  \vfill
  \vfill
  \vfill

\end{frame}

\subsection{The Sample Mean}

\begin{frame}
  \frametitle{Sample Mean}

  We have a bunch of measurements:
  \begin{eqnarray*}
    \bar{x} & = & x_1,~x_2,~x_3,\cdots,~x_n.
  \end{eqnarray*}
  Each measurement has a mean, $\mu$, and a standard deviation of
  $\sigma$. We assume that they are all independent of one another.
  
    \begin{definition}
      Given measurements the sample mean is 
      \begin{eqnarray*}
        \bar{x} & = & \frac{x_1+x_2+x_3+\cdots+x_n}{n}.
      \end{eqnarray*}
      The sample mean, $\bar{x}$,  has a mean of $\mu$, and it has a
      standard deviation of $\frac{\sigma}{\sqrt{n}}$.
    \end{definition}

    The sample mean is a random variable!

\end{frame}


\subsection{Examples}

\begin{frame}
  \frametitle{Example}

  The life time of batteries is normally distributed with a mean of
  19.8 hours and a standard deviation of 1.3 hours. What is the
  probability that a sample size of six batteries will give a sample
  mean that is more than 21.1 hours?

\end{frame}


\begin{frame}
  \frametitle{Example}

  You want to estimate the start up costs for restaurants in an
  area. If the mean start up costs are \$235,000 with a standard
  deviation of \$45,000 what is the probability that a sample of
  twenty restaurants will be between \$260,000 and \$205,000?

\end{frame}


\begin{frame}
  \frametitle{Example}

  You want to estimate the start up costs for restaurants in an
  area. If the mean start up costs are \$235,000 with a standard
  deviation of \$45,000 how many restaurants should you poll so that
  the probability to be within \$1,000 of the mean is ten percent?

\end{frame}


\begin{frame}
  \frametitle{Clicker Quiz}

  You want to estimate the start up costs for restaurants in an
  area. The mean start up costs are \$235,000 with a standard
  deviation of \$45,000. What is the probability that twenty samples
  will give a sample mean greater than \$245,000?


  \vfill

  \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
    A: 0.1611  & B: 0.4121  & C: 0.9999
  \end{tabular}

  \vfill
  \vfill
  \vfill

\end{frame}



% LocalWords:  Clarkson pausesection hideallsubsections
