
\lecture{Probability Rules}{probability-rules}
\section{Probability Rules}

\title{Probability Rules}
\subtitle{Putting it All Together}

\author{Kelly Black}
\institute{Clarkson University}
\date{1 February 2012}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[pausesection,hideothersubsections,sectionstyle=show/hide]
\end{frame}


\subsection{Clicker Quiz}


\begin{frame}
  \frametitle{Clicker Quiz}

  I have a bag with ten blue marbles and four red marbles. I remove
  two marbles without replacement. What is the probability that I pull
  out a red marble first and then a blue marble?

  \vspace{2em}

  \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l}
    A: $\frac{40}{196}$ & B: $\frac{40}{182}$ & C: $\frac{14}{140}$
  \end{tabular}

  \vfill

\end{frame}

\subsection{Probability Rules}

\begin{frame}{Probability Rules}

  \begin{definition}
    The multiplication rule:
    \begin{eqnarray*}
      p(A)\cdot p(B|A) & = & p(A \mathrm{~and~} B), \\
      p(B)\cdot p(A|B) & = & p(A \mathrm{~and~} B).
    \end{eqnarray*}

    The addition rule:
    \begin{eqnarray*}
      p(A \mathrm{~or~} B) & = & p(A) + p(B) - p(A \mathrm{~and~} B).
    \end{eqnarray*}

    The compliment rule:
    \begin{eqnarray*}
      p(\bar{A}) & = & 1 - p(A).
    \end{eqnarray*}

  \end{definition}

  \vfill
  
\end{frame}


\subsection{Examples}

\begin{frame}
  \frametitle{Example 1}

  \vfill

  A ball is hidden under one of three cups. You pay \$5.00 to find the
  ball. You get \$10.00 if you find it on your first guess. You get
  \$2.00 if you get it on your second guess. Otherwise you get
  nothing.

  \vfill

  What is the probability of getting some money back?

  \vfill

\end{frame}


\begin{frame}
  \frametitle{Clicker Quiz}
  
  \vfill

  A store sells skates. Thirty percent of the skates are used. Five
  percent of the new skates are defective. Twenty percent of the used
  skates are defective. You pick one set of skates out at random. What
  is the probability that the set is defective?

  \vfill

  \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
    A: 0.050  & B: .095  & C: .110 & D: .250
  \end{tabular}

  \vfill

\end{frame}


\begin{frame}{Example}

  \vfill

  A test can detect the presence of an abnormal gene that causes cystic
  fibrosis (CF). It is estimated that about 5\% of all people have the
  abnormal gene. The test is correct 90\% of the time for people who
  have the abnormal gene. About 2\% of the time it is positive for
  people who do not have the abnormal gene.

  \vfill

  If someone tests positive for the abnormal gene what is the
  probability that they have the abnormal gene?

  \vfill
  
\end{frame}

\begin{frame}
  \frametitle{Example}

  I have a bag with ten blue marbles and four red marbles. I remove
  two marbles without replacement. What is the probability that I pull
  out at least one blue marble?


  \vfill

\end{frame}



% LocalWords:  Clarkson pausesection hideallsubsections
