
\lecture{Independence And Mutual Exclusion}{indepence-and-mutual-exclusion}
\part{Independence And Mutual Exclusion}

\title{Independence And Mutual Exclusion}
\subtitle{Stuff We Already Did}

\author{Kelly Black}
\institute{Clarkson University}
\date{3 Feb 2012}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[pausesection,hideallsubsections,part=1]
\end{frame}


\section{Clicker Quiz}


\begin{frame}
  \frametitle{Clicker Quiz}

  If $p(A)=0.3$, $p(B)=0.4$, and $p(A\mathrm{~or~}B)=.37$ what is
  $p(A\mathrm{~and~}B)$?

  \vfill

  \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l}
    A: .11 & B: 12 & C: .33
  \end{tabular}

  \vfill
  \vfill
  \vfill

\end{frame}




\section{Example}

\begin{frame}
  \frametitle{Example 1}

  I flip a coin two times. What is the probability that I get a tails
  on the second flip?

  \vfill


\end{frame}


\section{Mutually Exclusive Events}

\begin{frame}
  \frametitle{Mutual Exclusion}

  \begin{definition}
    Two sets are \textit{mutually exclusive} if they have nothing in
    common.
  \end{definition}

\end{frame}


\begin{frame}{Example 2}
  
  I have a bag with six red marbles and two blue marbles. I draw two
  without replacement. What is the probability that the second marble
  is red?

  \vfill

\end{frame}

\section{Independent Events}

\begin{frame}
  \frametitle{Independence}

  \begin{definition}
    Two events are independent if 
    \begin{eqnarray*}
      p(A|B) & = & p(A).
    \end{eqnarray*}

    This is equivalent to requiring that
    \begin{eqnarray*}
      p(B|A) & = & p(B).
    \end{eqnarray*}

  \end{definition}

\end{frame}



\begin{frame}{Example 3}

  Suppose that $p(A)=0.6$, $p(B)=0.3$, and $p(A\mathrm{~and~}B)=.2$
  Are they independent?

  \vfill

  
\end{frame}



\begin{frame}{Example 4}

  Suppose that $p(A)=0.6$, $p(B)=0.3$, and $p(A\mathrm{~and~}B)=.18$
  Are they independent?

  \vfill

  
\end{frame}


\begin{frame}{Example 5}
  An industrial process is established to produce items. About five
  percent of the items are defective under this process. Each item is
  inspected, and fifteen percent of the defective items are not caught
  in the inspection. About three percent of the good items are
  classified as defective.

  \begin{itemize}
  \item What percentage of the items that are sold are defective?
  \item I pick a part at random. That part is classified as
    defective. What is the probability that it is actually defective?
  \end{itemize}

  \vfill

\end{frame}


\begin{frame}{Example 6}

  A brokerage firm has a system that can execute a sell request in
  within 0.08 seconds about 85\% of the time. Of the requests that
  occur within 0.08 seconds about 75\% meet the projected cost
  level. Of the requests that do not occur within 0.08 seconds about
  65\% meet the projected cost level.

  What is the proportion of requests that will meet the projected cost
  levels?

  \vfill
  
\end{frame}




% LocalWords:  Clarkson pausesection hideallsubsections
