
\lecture{Hypothesis Testing with the Classical Approach}{hypothesis-testing-classical}
\part{Hypothesis Testing with the Classical Approach}

\title{Hypothesis Testing Again}
\subtitle{The ``Classical'' Approach}

\author{Kelly Black}
\institute{Clarkson University}
\date{14 March 2012}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[pausesection,hideallsubsections,part=1]
\end{frame}


\section{Clicker Quiz}


\begin{frame}
  \frametitle{Clicker Quiz}

  \iftoggle{clicker}{%

    I read in the \textit{Integrator} that the mean height of Clarkson
    students is 180.0 cm with a standard deviation of 14.0 cm. I think
    that this is too high and sample twenty students chosen at random
    to find a sample mean of 175.1 cm. Are my suspicions correct? (Use
    a 5\% significance level.)

    \vfill

    \begin{tabular}{l@{\hspace{3em}}l}
      A: & There is sufficient evidence to reject $H_0$.  \\
      B: & There is not sufficient evidence to reject $H_0$.  
    \end{tabular}

    \vfill
    \vfill
    \vfill

    

  }


\end{frame}

\section{Classical Approach}

\begin{frame}
  \frametitle{What if....}


  I read in the \textit{Integrator} that the mean height of Clarkson
  students is 180.0 cm with a standard deviation of 14 cm. I think
  that this is too high and sample twenty students chosen at random to
  find a sample mean of \textbf{174.7} cm. Are my suspicions correct?
  (Use a 5\% significance level.)

    \vfill
    

\end{frame}


\begin{frame}{Result}

  We have sufficient evidence to reject $H_0$ at the 5\% significance
  level using a normal distribution with twenty samples and a standard
  deviation of 14.0cm.
  
\end{frame}





\begin{frame}{Clicker Quiz}

  \iftoggle{clicker}{%

    I see a publication that says that the mean house value in an area
    is \$215,000 with a standard deviation of \$35,000. I think that
    this is too low.  Suppose that I take a random sample of fifteen
    homes and get a sample mean of \$232,000.

    \vfill 

    What is the $z$-statistic?

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: -1.881  & B: -1.644  & C: 1.644 & D: 1.881
    \end{tabular}

    \vfill
    \vfill
    \vfill

  }

\end{frame}

\begin{frame}{Result}

  We have sufficient evidence to reject $H_0$ at the 5\% significance
  level using a normal distribution with fifteen samples and a standard
  deviation of \$35,000.
  
\end{frame}


\begin{frame}{Clicker Quiz}

  \iftoggle{clicker}{%

    \begin{eqnarray*}
      \begin{array}{lrcl}
        H_0: & \mu & = & 800 \\
        H_a: & \mu & \neq & 800
      \end{array}
    \end{eqnarray*}

    \begin{eqnarray*}
      n & = & 30, \\
      \bar{x} & = & 758,\\
      \sigma & = & 118, \\
      \alpha & = & 0.05.
    \end{eqnarray*}

    \vfill 

    What is the $z$-statistic?

    \begin{tabular}{l@{\hspace{3em}}l@{\hspace{3em}}l@{\hspace{3em}}l}
      A: -1.960  & B: -1.950  & C: -0.065 & D: 1.960
    \end{tabular}

    \vfill
    \vfill
    \vfill

  }

\end{frame}

\begin{frame}{Result}

  We \textbf{do not} have sufficient evidence to reject $H_0$ at the
  5\% significance level using a normal distribution with thirty
  samples and a standard deviation of 118.
  
\end{frame}




% LocalWords:  Clarkson pausesection hideallsubsections
